\textls[-0]{\thetitle}

### 2023-03-22 casestudies.tex
ImpCP

The code is organized into four families.
A base family, \lsti{Imp}, defines via \lsti{FInductive} the abstract
syntax of a while-language with pure expressions and impure statements.
The semantics is given by an interpreter defined as a CEK-style abstract
machine~\cite{felleisen1986control} and parameterized by a fuel value.
Family \lsti{Imp} defines the interpreter via \lsti{FRecursion}.
It contains \textasciitilde 200~LOC.

A second family, \lsti{ImpGAI}, extends \lsti{Imp} and uses
\textasciitilde 550~LOC.
It exports a generic framework for deriving
abstract interpreters with partial-correctness guarantees.
Soundness of the abstract interpreter, \lsti{analyze},
is stated with respect to the interpreter, \lsti{eval}, inherited from \lsti{Imp}.
The theorem says that the abstraction relation \lsti{RState} over a
concrete state~\lsti{S} and an abstract state~\lsti{absS} is preserved
by the analysis:

@@

The other two families then extend \lsti{ImpGAI}.
Family \lsti{ImpTI} (\textasciitilde200~LOC) is an abstract
interpreter that does type inference~\cite{cousot1997types}.
Family \lsti{ImpCP} (\textasciitilde300~LOC) extends the abstract syntax
with natural number arithmetic,
and further binds the generic abstract interpreter to perform constant propagation.

Our implementation of family polymorphism is compatible with Coq's
program extraction feature.
We extract the two verified abstract interpreters to OCaml
and put them to test over simple queries. %\rD{What were the results of testing the extracted abstract interpreters? I don't expect much in the way of performance, but did they return the correct results? What sorts of benchmarks did you use?}

We note that our plugin cannot yet allow the expression of two derived abstract
interpreters nested within the same family.
Work on nested inheritance~\cite{ncm2004,zm2017} points to a direction
to further increase the expressive power of our language design.

If not using \Lang, we can construct \lsti{Imp}, \lsti{ImpGAI} and \lsti{ImpTI} in a similar way---but \lsti{Imp} is now a module instead of a family, \lsti{ImpGAI} is now a functor instead of a family with overridable fields, and \lsti{ImpTI} is a concrete instantiation of the functor \lsti{ImpGAI}. However, \lsti{ImpCP} cannot reuse the above framework as it extends syntax with natural numbers. Thus \lsti{ImpCP} will require repeating a large amount of definitions, and it alone will needs about 900~LOCs.

>>>>

The code is organized into four families.
A base family \lsti{Imp} (\textasciitilde 200~LOC) defines via \lsti{FInductive} the abstract
syntax of a while-language with pure expressions and impure statements.
The semantics is given by an interpreter defined as a CEK-style abstract
machine~\cite{felleisen1986control} and parameterized by a fuel value.
Family \lsti{Imp} defines the interpreter via \lsti{FRecursion}.

A second family \lsti{ImpGAI} (\textasciitilde 550~LOC) extends \lsti{Imp}.
It exports a generic framework for deriving
abstract interpreters with partial-correctness guarantees.
Soundness of the abstract interpreter, \lsti{analyze},
is stated with respect to the interpreter, \lsti{eval}, inherited from \lsti{Imp}.
The theorem says that the abstraction relation \lsti{RState} over a
concrete state~\lsti{S} and an abstract state~\lsti{absS} is preserved
by the analysis:

@@

The next two families then extend \lsti{ImpGAI}.
Family \lsti{ImpTI} (\textasciitilde200~LOC) is an abstract
interpreter that does type inference~\cite{cousot1997types}.
Family \lsti{ImpCP} (\textasciitilde300~LOC) extends the abstract syntax
with natural-number arithmetic,
and further binds the generic abstract interpreter to perform constant propagation.
%
An alternative to \Lang would be to define \lsti{ImpGAI} as a functor parameterized
by the abstract domain and such. But it would prevent \lsti{ImpCP} from reusing
the code and proofs in \lsti{ImpGAI}, because \lsti{ImpCP} involves extending
inductive types with new constructors.

Our implementation of family polymorphism is compatible with Coq's
program extraction feature.
We extract the two verified abstract interpreters to OCaml.
Testing the extracted program over simple queries returns expected results.
%\rD{What were the results of testing the extracted abstract interpreters? I don't expect much in the way of performance, but did they return the correct results? What sorts of benchmarks did you use?}

We note that \Lang cannot yet allow the expression of two derived abstract
interpreters nested within the same family.
Work on nested inheritance~\cite{ncm2004,zm2017} points to a direction
to further increase the expressive power of our language design.

%If not using \Lang, we can construct \lsti{Imp}, \lsti{ImpGAI} and \lsti{ImpTI} in a similar way---but \lsti{Imp} is now a module instead of a family, \lsti{ImpGAI} is now a functor instead of a family with overridable fields, and \lsti{ImpTI} is a concrete instantiation of the functor \lsti{ImpGAI}. However, \lsti{ImpCP} cannot reuse the above framework as it extends syntax with natural numbers. Thus \lsti{ImpCP} will require repeating a large amount of definitions, and it alone will needs about 900~LOCs.
